### 2020-02-02 PLACES_2020/abstract.tex

\begin{abstract}
Advancements in mobile device compute power have made interactive web
applications possible, allowing the web browser to render contents dynamically
and support low-latency communication with the server.
This comes at a cost to the developer, who now needs to reason more about the
communication correctness of their application as they support more complex
communication patterns.

>>>>

\begin{abstract}
Advancements in mobile device compute power have made interactive web
applications possible, allowing the web browser to render content dynamically
and support low-latency communication with the server.
This comes at a cost to the developer, who now needs to reason more about the
communication correctness of their application as they support more complex
communication patterns.

### 2020-02-02 PLACES_2020/conclusion.tex

\section{Conclusion and Future Works}
We have presented a MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications that conform to a
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

>>>>

\section{Conclusion and Future Work}
We have presented a MPST-based framework for developing full-stack interactive
TypeScript applications with WebSocket communications that conform to a
protocol, statically providing linear channel usage guarantees and affine
channel usage guarantees for back-end and front-end targets respectively.

### 2020-02-02 PLACES_2020/intro.tex
Address comments from Francisco

% Example: noughts and crosses
Consider a classic turn-based board game of \textit{Noughts and Crosses}
between two players.
Both players, identified by either noughts or crosses respectively, take
turns to place a mark on an unoccupied cell of a 3-by-3 grid until one player
wins (when their markers form one straight line on the board) or a stalemate is
reached (when all cells are occupied).
A web-based implementation may involve players connected to a game server via
WebSocket connections.
The players interact with the game from their web browser, which shows a
\textit{single-page application} (SPA) of the game client written in a popular
framework like \textit{React.js} \cite{React}.
SPAs feature a single HTML page and dynamically rendered content via JavaScript
in the browser.
Players take turns to make a move on the game board, which sends a message to
the server, and the server implements the game logic to progress the game
forward until a result (either a win/loss or draw) can be declared, where
either the move of the other player or the game result is sent to players.

Whilst WebSockets make this web-based implementation possible, it introduces
the developer to a new family of communication errors, even for this simple
game.
In addition to the usual testing for game logic correctness, the
developer needs to test against \textit{deadlocks} (e.g. both players waiting
for each other to make a move at the same time) and \textit{communication
 mismatches} (e.g. player 1 sending a boolean to the game server instead of
the board coordinates).
The complexity of these errors correlate to the
complexity of related tests required, scale with the complexity of
communication patterns involved.

>>>>

% Example: noughts and crosses
Consider a classic turn-based board game of \textit{Noughts and Crosses}
between two players.
Both players, identified by either noughts or crosses respectively, take
turns to place a mark on an unoccupied cell of a 3-by-3 grid until one player
wins (when their markers form one straight line on the board) or a stalemate is
reached (when all cells are occupied).
A web-based implementation may involve players connected to a game server via
WebSocket connections.
The players interact with the game from their web browser, which shows a
\textit{single-page application} (SPA) of the game client written in a popular
framework like \textit{React.js} \cite{React}.
SPAs feature a single HTML page and dynamically render content via JavaScript
in the browser.
Players take turns to make a move on the game board, which sends a message to
the server, and the server implements the game logic to progress the game
forward until a result (either a win/loss or draw) can be declared, where
either the move of the other player or the game result is sent to players.

Whilst WebSockets make this web-based implementation possible, it introduces
the developer to a new family of communication errors, even for this simple
game.
In addition to the usual testing for game logic correctness, the
developer needs to test against \textit{deadlocks} (e.g. both players waiting
for each other to make a move at the same time) and \textit{communication
 mismatches} (e.g. player 1 sending a boolean to the game server instead of
the board coordinates).
The complexity of these errors correlate to the
complexity of the required tests and scale with the complexity of
communication patterns involved.

### 2020-02-02 PLACES_2020/related.tex

\section{Related Works}
Native language support for first-class linear channel resources \cite{ATS} and
code generation are two main approaches for incorporating MPST into application
development.
The latter closely relates to our proposal;
we highlight two areas of existing works under this approach that motivate our
design choice.

>>>>

\section{Related Work}
Native language support for first-class linear channel resources \cite{ATS} and
code generation are two main approaches for incorporating MPST into application
development.
The latter closely relates to our proposal;
we highlight two areas of existing work under this approach that motivate our
design choice.

### 2020-02-02 interim/sections/background/scribble.tex

\cite{Hybrid2016} also identifies properties of EFSMs for well-formed protocol specifications: \textbf{(1)} there is exactly one initial state, \textbf{(2)} there is at most one terminal state, and \textbf{(3)} every non-terminal state is either an input state or output state. There are further restrictions on input and output states, namely these states must only send to or receive from the same role. We formalise these properties in Figure \ref{fig:efsmstates}.

>>>>

We adopt the properties of EFSMs for well-formed protocol specifications from \cite{Hybrid2016} \textbf{(1)} there is exactly one initial state, \textbf{(2)} there is at most one terminal state, and \textbf{(3)} every non-terminal state is either an input state or output state. There are further restrictions on input and output states, namely these states must only send to or receive from the same role. We formalise these properties in Figure \ref{fig:efsmstates}.

### 2020-02-02 interim/sections/introduction.tex

Modern interactive web applications aim to provide a highly responsive user experience by minimising the communication latency between clients and servers. 
Whilst the HTTP request-response model is sufficient for retrieving static assets, applying the same stateless communication approach for interactive use cases (such as a multiplayer game) introduces undesirable performance overhead from having to frequently set up new connections for client-server interactions. 
Developers have since adopted other communication transport abstractions over HTTP connections such as the WebSockets protocol \cite{WebSocketRFC} to enjoy low-latency full-duplex client-server communication in their applications over a single persistent connection. 
Enabling more complex communication patterns caters for more interactive use cases (such as real-time multiplayer games), but introduces additional concerns to the developer with respect to implementation correctness.

@@

We present a workflow for developing type-safe interactive SPAs motivated by the MPST framework: \textbf{(1)} An endpoint API code generation workflow targeting TypeScript-based web applications for multiparty sessions; \textbf{(2)} An encoding of session types in server-side TypeScript that enforces static linearity; and \textbf{(3)} An encoding of session types in browser-side TypeScript using the React framework that guarantees affine usage of communication channels.

>>>>

Modern interactive web applications aim to provide a highly responsive user experience by minimising the communication latency between clients and servers. 
Whilst the HTTP request-response model is sufficient for retrieving static assets, applying the same stateless communication approach for interactive use cases (such as multiplayer games) introduces undesirable performance overhead from having to frequently set up new connections for client-server interactions. 
Developers have since adopted other communication transport abstractions over HTTP connections such as the WebSockets protocol \cite{WebSocketRFC} to enjoy low-latency full-duplex client-server communication in their applications over a single persistent connection. 
Enabling more complex communication patterns caters for more interactive use cases, but introduces additional concerns to the developer with respect to implementation correctness.

@@

\paragraph{Contributions} We present a workflow for developing type-safe interactive SPAs motivated by the MPST framework:

\begin{enumerate}
\item An endpoint API code generation workflow targeting TypeScript-based web applications for multiparty sessions;
\item An encoding of session types in server-side TypeScript that enforces static linearity;
\item An encoding of session types in browser-side TypeScript using the React framework that guarantees affine usage of communication channels.
\end{enumerate}
